---
title: "502 homework2"
author: "Songran Wang"
date: "2024-10-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1

```{r}
set.seed(123)
samples_A <- rnorm(30, mean = 0, sd = sqrt(1))
samples_B <- rnorm(10, mean = 0, sd = sqrt(4))
```

## (a)

```{r}
t_stat_two<-function(yA,yB){
  nA<-length(yA)
  nB<-length(yB)
  sp<-sqrt(((nA-1)*sd(yA)**2+(nB-1)*sd(yB)**2)/(nA+nB-2))
  denominator<-sp*sqrt((1/nA)+(1/nB))
  numerator<-(mean(yA)-mean(yB))
  return (numerator/denominator)
}

observed_t_stat <- t_stat_two(samples_A, samples_B)
observed_t_stat
```

The observed two-sample t-stat is $-1.895$

## (b)

```{r}
set.seed(123)
combined_samples <- c(samples_A, samples_B)# sample_A+sample_B
simulated_t_stats <- numeric(10000)#store samples t stats values
for (i in 1:10000)
{
  permuted_samples=sample(combined_samples,replace = TRUE)
  permuted_A <- permuted_samples[1:30]
  permuted_B <- permuted_samples[31:40]
  simulated_t_stats[i] <- t_stat_two(permuted_A, permuted_B)
}
p_value <- mean(abs(simulated_t_stats) >= abs(observed_t_stat))
cat("P-value of the permutation test:", p_value, "\n")
```

$H_0: \mu_A=\mu_B$

because 0.0707\>0.05, we could not reject the null hypothesis.

## (c)

### i

$t\left(\mathbf{Y}_A, \mathbf{Y}_B\right)=\frac{\bar{Y}_B-\bar{Y}_A}{s_p \sqrt{\frac{1}{n_A}+\frac{1}{n_B}}} \sim t_{n_A+n_B-2}$

degrees of freedom is 38

### ii

```{r}
critical_value <- qt(1 - 0.05/2, 38)
cat('the critical value is:',critical_value)
```

the critical value is 2.024

### iii

```{r}
p_value_t_test <- 2 * (1 - pt(abs(observed_t_stat), 38))
cat('p value of the two samples t test:',p_value_t_test)
```

the p value is 0.066, so we cannot reject the null hypothesis.

# 2

## (a)

$t_w\left(\boldsymbol{y}_A, \boldsymbol{y}_B\right)=\frac{\bar{y}_B-\bar{y}_A}{\sqrt{\frac{s_A^2}{n_A}+\frac{s_B^2}{n_B}}}$

```{r}
welch_t_stat=(mean(samples_A)-mean(samples_B))/sqrt(var(samples_A)/length(samples_A)+var(samples_B)/length(samples_B))
welch_t_stat
```

so the observed test statistic is -1.826

## (b)

$t_w\left(\boldsymbol{y}_A, \boldsymbol{y}_B\right)=\frac{\bar{y}_B-\bar{y}_A}{\sqrt{\frac{s_A^2}{n_A}+\frac{s_B^2}{n_B}}}$, this stat test is a $t_{\nu_w}$-distribution, where $\nu_w=\frac{\left(\frac{s_A^2}{n_A}+\frac{s_B^2}{n_B}\right)^2}{\frac{1}{n_A-1}\left(\frac{s_A^2}{n_A}\right)^2+\frac{1}{n_B-1}\left(\frac{s_B^2}{n_B}\right)^2}$

```{r}
v_w=(var(samples_A)/length(samples_A)+var(samples_B)/length(samples_B))^2/(1/(length(samples_A)-1)*(var(samples_A)/length(samples_A))^2+1/(length(samples_B)-1)*(var(samples_B)/length(samples_B))^2)
v_w
```

so the freedom degree is 14.565

## (c)

```{r}
critical_value <- qt(1 - 0.05/2, v_w)
cat('the critical value is:',critical_value)
```

so the critical value is 2.137

## (d)

```{r}
p_value_t_test <- 2 * (1 - pt(abs(welch_t_stat), v_w))
cat('p value of the two samples t test:',p_value_t_test)
```

the p value is 0.088, so we cannot reject the null hypothesis.

## (e)

In the Problem 1, it is assumed that the two groups have the same variance, however the Welch's t-test is not. Two statistics in each test are different as well.

# 3

## (a)

```{r}
set.seed(123)



# i.  long-tailed distribution(normal cauchy)
long_tailed_samples <- rcauchy(1000)
qqnorm(long_tailed_samples, main = "Long-Tailed Distribution (normal cauchy)")
qqline(long_tailed_samples, col = "red")

# ii.  short-tailed distribution(normal)
short_tailed_samples <- rnorm(1000)
qqnorm(short_tailed_samples, main = "Short-Tailed Distribution (normal)")
qqline(short_tailed_samples, col = "red")

# iii.  left-skewed distribution(beta distribution(6,2))
left_skewed_samples <- rbeta(1000, 6, 2)
qqnorm(left_skewed_samples, main = "Left-Skewed Distribution (beta (6,2))")
qqline(left_skewed_samples, col = "red")

# iv. right-skewed distribution(Chi-squared(3))
right_skewed_samples <- rchisq(1000, df=3)
qqnorm(right_skewed_samples, main = "Right-Skewed Distribution (Chi-squared(3))")
qqline(right_skewed_samples, col = "red")

```

## (b)

1.  Top left corner: There are some points deviate the line at the end, which seems that this distribution is long-tailed.
2.  Top right corner: The middle of points line is curved, with some deviation at both ends, which means that this distribution is right-skewed.
3.  Bottom left corner: Points at both ends see extreme deviations, so the distribution is long-tailed. Also it seems it is left-skewed.
4.  Bottom right corner: Points almost follow the line, showing that this distribution is nearly normal.

# 4

```{r}
feather <- data.frame(
  Typical = c(-0.255, -0.213, -0.190, -0.185, -0.045, -0.025, -0.015, 
              0.003, 0.015, 0.020, 0.023, 0.040, 0.040, 0.050, 0.055, 0.058),
  Odd = c(-0.324, -0.185, -0.299, -0.144, -0.027, -0.039, -0.264, 
          -0.077, -0.017, -0.169, -0.096, -0.330, -0.346, -0.191, -0.128, -0.182)
)
```

## (a)

the difference between two paired groups is $d_i=$ $x_{1 i}-x_{2 i}$

$\bar{d}=\frac{1}{n} \sum_{i=1}^n d_i$

$s_d=\sqrt{\frac{\sum_{i=1}^n\left(d_i-\bar{d}\right)^2}{n-1}}$

$t=\frac{\bar{d}}{s_d / \sqrt{n}}\sim t(n-1)$

```{r}
t_test_paired <- t.test(feather$Typical, feather$Odd, paired = TRUE)
t_test_paired
```

so the test statistic is $4.0647$, and it follow $t(15)$ distribution.

## (b)

the non-centrality parameter is $\frac{\delta}{S \sqrt{\frac{1}{n}}}$, where $\delta=0.05$

```{r}
typical <- feather$Typical
odd <- feather$Odd

n <- length(typical)

s<-var(typical-odd)

ncp=0.05/(sqrt(s)*sqrt(1/n))
ncp
```

the non-centrality parameter is $1.482$

## (c)

$$
\begin{aligned}
\operatorname{Power}(\delta) & =\operatorname{Pr}\left(\text { reject } H_0 \mid \mu=\delta\right) \\
& =\operatorname{Pr}\left(\left|t\right| \geq t_{1-\alpha / 2, n-1} \mid \mu=\delta\right)\\
& =\operatorname{Pr}\left(t>t_{1-\alpha / 2, n-1}\right)+\operatorname{Pr}\left(t<-t_{1-\alpha / 2, n-1}\right) \\
& =\left[1-\operatorname{Pr}\left(t<t_{1-\alpha / 2, n-1}\right)\right]+\operatorname{Pr}\left(t<-t_{1-\alpha / 2, n-1}\right)
\end{aligned}
$$

where $\delta=0.05$

```{r}
t.crit <-qt( 1-0.05/2 , n-1 )
t.gamma=ncp
t.power<-1-pt(t.crit,n-1,ncp=t.gamma )+pt(-t.crit,n-1,ncp=t.gamma )
t.power
```

the power is 28.4%

## (d)

```{r}
n<-seq(20,150)
t.crit<-qt(1-0.05/2,n-1)
t.gamma=0.05/(sqrt(s)*sqrt(1/n))
t.power<-1-pt(t.crit,n-1,ncp=t.gamma )+pt(-t.crit,n-1,ncp=t.gamma )
plot(n, t.power, xlab = "Sample Size (n)", ylab = "Power (t.power)", 
     main = "Scatterplot of t.power vs. Sample Size", 
     pch = 19)
index <- which(t.power >= 0.9)[1]
points(n[index], t.power[index], col = "red", pch = 19, cex = 1.5) 
text(n[index],t.power[index], labels = paste("(", n[index], ",", t.power[index], ")", sep = ""), pos = 4, col = "red")
```

when n=79 for each group, the power is over 90%
